Assessment R Markdown
========================================================

## This is a R markdown for the Assessment (Peer Reviewed) for the Practical Machine Learning Course from the John Hopkinson University

### **Author** - Mohit Sewak - *mohitsewak@gmail.com*

#### Built on R version 2.15.2, using R Studio Version 0.98.507, Posted on GitHub Repository [mohitsewak/RMachineLearning] (http://mohitsewak.github.io/RMachineLearinng/) using gw-pages.


Loading required libraries (Results Hidden, Warnings OFF, Messages OFF)

```{r libraries, results='hide',warning=FALSE,message=FALSE}
library(lattice)
library(ggplot2)
library(graphics)
library(Hmisc)
library(e1071)
library(caret)

```

Reading (& caching) Files.

```{r reading_files, cache=TRUE}
training = read.csv("pml-training.csv")
testing =  read.csv("pml-testing.csv")
```


Creating summary and viewing the top few records, and seeing paired scatter plots in the training file (Results Hidden, Output Cached).


```{r data_summary, results="hide", cache=TRUE}
summary(training)
head(training)
```

Since the data size, and number of feature vectors are huge, so reducing both the data size for plots and also the number of columns in a set of paired visualization

```{r pair_plots, cache=TRUE,fig.width=7, fig.height=6}
plot(training)
pairs(training[1:10000,1:10])
pairs(training[1:10000,11:20])
pairs(training[1:10000,21:30])
```

Since most of the columns have no data, or predictive power, it might not be conducive to use them as-is.
<p> Conducting the following prepocessing operations:</p>
- Removing Near Zero Value Predictors
- Principal Component Axis (PCA) rotation to optimize/ reduce dimensions in the data set 
- (Nt: Variable 160 = target variable named "classe"")

```{r preProcess_NZV, cache=TRUE}
nzv <- nearZeroVar(training[,-160])
dim(training)
nzv_training <- training[,-nzv]
dim(nzv_training)
preProcess(nzv_training[,-160], method="pca", thresh=0.8)
```

Finding numric variables, as PCA can be applied only on numeric variables

```{r preProcess_PCA, cache=TRUE}
numericvars <- NULL
for (Var in names(nzv_training)) {
  if(class(nzv_training[,Var]) == 'integer' | class(nzv_training[,Var]) == 'numeric') {
    numericvars$Var <- c(numericvars,Var)
    numericvars$IsNumeric <- c(numericvars,TRUE)
    }
  else {
    numericvars$Var <- c(numericvars,Var)
    numericvars$IsNumeric <- c(numericvars,FALSE)
    }
  }

numericvars
```


